---
title: "R-code for 'TITLE OF MANUSCRIPT'"
author: "NAMES OF AUTHORS"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, tidy=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document provides all the `R code` used in our paper. Both the Rmarkdown file and the data can be downloaded from the accompanying GitHub repository on (https://github.com/TonyRinaud/malarone_tolerability) as a zip archive containing all the files. We recommend to download or clone this [GitHub repository](https://github.com/TonyRinaud/malarone_tolerability) in order to access the documentation together with all the files that are needed to repeat analyses shown in this document. Just click on the link above and then on the green box `Clone or download`. In order to function properly, the same structure of folders must be kept. If you have any questions, don't hesitate to contact tony.rinaud[at]uni-bielefeld.de or m.ottensmann[at]uni-bielefeld.de

In order to repeat analyses presented in this document, the following packages are required. 

```{r packages, message=F, warning=F}
## Load packages
## -----------------------------------------------------------------------------
easypackages::libraries(
  "tidyverse", "lme4", "chron", "patchwork",
  "magrittr", "readr", "EnvStats", "sjPlot", "egg") 
```

## Loading raw datasets

* `Malarone experiment`

Experimental study comprising data from 299 nestlings that were either treated with Malarone or water. An additional subset comprises individuals that received no treatment and were therefore only exposed to repeated routine sampling and handling. 


```{r echo = TRUE, warning=FALSE}
## Read and format raw data: (saved raw data with "." as decimal operator)
## -----------------------------------------------------------------------------
malarone.data <- readr::read_delim(
  file = 'data/malarone_tolerability.csv', delim = ",", col_types = "fffffffinnnnnnnnnnnffff") %>% 
  mutate(Treatment = factor(Treatment, levels=c("Control","Water","Malarone"))) %>% 
  ## Scaling continuous variables
  mutate(delta_body_condition = scale(delta_body_condition)) %>% 
  mutate(body_condition.1 = scale(body_condition.1)) %>% 
  mutate(body_condition.2 = scale(body_condition.2)) %>% 
  mutate(growth_rate = scale(growth_rate))
```

Summarise dataset:

```{r}
table(malarone.data$Year)
table(malarone.data$Treatment)
quantile(malarone.data$avg_age)
quantile(malarone.data$days_resample)
```


* `Bloodchemistry data`

Dataset on blood chemistry values for 11 individuals of the `Malarone experiment` prior and after treatment. Additionally, blood chemistry values for another 60 individuals that were not part of the `Malarone experiment`.


```{r}
bloodchem.data <- readr::read_csv(file = "data/bloodchemistry_data.csv") %>% 
  mutate(Ring = as.character(Ring)) %>% 
  mutate(Treatment = factor(Treatment,
                            levels = c("Pre-Control", "Post-Control", "Pre-Mal", "Post-Mal"))) %>% 
  mutate(doy = lubridate::yday(Date)) 
names(bloodchem.data)[names(bloodchem.data) == "Lprev"] <- "Leucocytozoon"

## Some gGT values are at boundaries, set those to 5
## -----------------------------------------------------------------------------
bloodchem.data$GGT[bloodchem.data$GGT == "<5"] <- 5
bloodchem.data$GGT <- as.numeric(bloodchem.data$GGT)

## Summarise sample sizes
## -----------------------------------------------------------------------------
pander::pandoc.table(table(bloodchem.data$Treatment))
```

## Analyses
### `Experimental data`

**We should consider variable intervals between sampling events! Age_ind_days does not account for this, nor is it clear if treatment effects would need time to show up** 

#### Treatment effects on body condition 

First, visualising raw data 

```{r} 
g1_t1 <- ggplot(malarone.data)+
  aes(Treatment, body_condition.1) +
  geom_hline(yintercept = mean(malarone.data$body_condition.1, na.rm=T), linetype = "dashed") +
  geom_boxplot(color ="black", outlier.shape = NA, size=.5, width=.3, na.rm = T)+
  geom_jitter(color='lightgrey', width = 0.2, height = 0, na.rm = T) +
  stat_n_text(na.rm = T, y.pos = -Inf, vjust = -1)+
  labs(x="", y="Std. body condition index")+
  stat_summary(fun=mean, colour = c("#1F78B4", "#33A02C", "#FF7F00"), geom="errorbar", 
               size=1.5 , width=0.05, show.legend = FALSE, 
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x), na.rm = T) +
  stat_summary(fun=mean, colour = c("#1F78B4", "#33A02C", "#FF7F00"), geom="point", 
               size=5 ,show.legend = FALSE, na.rm = T) + 
  theme_article() +
  theme(legend.position="none",
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm")) +
  scale_y_continuous(limits = c(-4.5, 4.75), breaks = seq(-4, 4, 1)) +
  ggtitle("1st sampling")


g1_t2 <- ggplot(malarone.data)+
  aes(Treatment, body_condition.2)+
  geom_hline(yintercept = mean(malarone.data$body_condition.2, na.rm=T), linetype = "dashed")+
  geom_boxplot(color ="black", outlier.shape = NA, size=.5, width=.3, na.rm = T)+
  geom_jitter(color='lightgrey', width = 0.2, height = 0, na.rm = T) +
  stat_n_text(na.rm = T, y.pos = -Inf, vjust = -1)+
  labs(x="", y="Std. body condition index")+
  stat_summary(fun=mean, colour = c("#1F78B4", "#33A02C", "#FF7F00"), geom="errorbar", 
               size=1.5 , width=0.05, show.legend = FALSE, 
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x), na.rm = T)+
  stat_summary(fun=mean, colour = c("#1F78B4", "#33A02C", "#FF7F00"), geom="point", 
               size=5 ,show.legend = FALSE, na.rm = T)+ 
  theme_article() +
  theme(legend.position="none",
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm")) +
  scale_y_continuous(limits = c(-4.5, 4.75), breaks = seq(-4, 4, 1)) +
  ggtitle("2nd sampling")

g1_delta <- ggplot(malarone.data)+
  aes(Treatment,delta_body_condition)+
  geom_hline(yintercept = mean(malarone.data$delta_body_condition, na.rm=T), linetype = "dashed")+
  geom_boxplot(color ="black", outlier.shape = NA, size=.5, width=.3, na.rm = T)+
  geom_jitter(color='lightgrey', width = 0.2, height = 0, na.rm = T) +
  stat_n_text(na.rm = T, y.pos = -Inf, vjust = -1)+
  labs(x="", y="\U0394 Std. body condition index")+
  stat_summary(fun=mean, colour = c("#1F78B4", "#33A02C", "#FF7F00"), geom="errorbar", 
               size=1.5 , width=0.05, show.legend = FALSE, 
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x), na.rm = T)+
  stat_summary(fun=mean, colour = c("#1F78B4", "#33A02C", "#FF7F00"), geom="point", 
               size=5 ,show.legend = FALSE, na.rm = T)+ 
  theme_article() +
  theme(legend.position="none",
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm")) +
  scale_y_continuous(limits = c(-4.5, 4.75), breaks = seq(-4, 4, 1)) +
  ggtitle("\U0394 BCI")

g1_t1 + g1_t2 + g1_delta + plot_annotation(tag_levels = 'A')& 
  theme(plot.tag = element_text(size = 14))
```

As shown above, there was some heterogeneity among treatment in the mean body condition index before treatment. Therefore, we used the change in body condition ($\Delta (2nd~sample - 1st~sample$)  

```{r}
## Run LMM
## -----------------------------------------------------------------------------
model.bci <- lmerTest::lmer(delta_body_condition ~ Treatment + avg_age + days_resample + Year + (1|Nest),
                  data = malarone.data)

## Summarise model
## -----------------------------------------------------------------------------
summary.model <- summary(model.bci)[["coefficients"]]
pander::pandoc.table(summary.model)

## Check model assumptions
## -----------------------------------------------------------------------------
performance::check_model(model.bci)
performance::model_performance(model.bci)

## Predicted values for each treatment
## -----------------------------------------------------------------------------
pred.bci <- get_model_data(model.bci, type = 'pred', terms = "Treatment") %>% 
  mutate(x = factor(x = c("Control","Water","Malarone"), levels = c("Control","Water","Malarone")))

## add sample size to labels
## -----------------------------------------------------------------------------
n <- table(malarone.data$Treatment[!is.na(malarone.data$delta_body_condition)])
pred.bci <- pred.bci %>% 
  mutate(x = paste0(x,"\nn=",n))

## Plot:
## 1.) predictions and corresponding error margins
## 2.) indicate significant differences based on tukey contrasts
## 3.) Jitter not fitting as corrected for considerable among-year variation
## -----------------------------------------------------------------------------
g1_pred_delta <- ggplot(pred.bci, aes(x = x, y = predicted)) +
  # geom_hline(yintercept = mean(pred.bci$predicted, na.rm = T), linetype = "dashed")+
  labs(x = "", y = "\U0394 body condition\n(z-transformed)")+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), colour = 'black', size = 1.5, width = .1) +
  geom_errorbar(aes(ymin = predicted - std.error, ymax = predicted + std.error) ,
                size = 8, colour = c("#1F78B4", "#33A02C", "#FF7F00"), width = 0) +
  geom_point(color ="black", size = 2) +
  scale_y_continuous(breaks = seq(0, 0.6, 0.1), limits = c(-0.06, 0.64)) +
  theme_article(base_size = 12) + 
  theme(legend.position="none",
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"))
```


#### Growth rate between control and malarone treated nestlings


Visualising the raw data

```{r} 
g2_water <- ggplot(malarone.data, aes(Treatment, growth_rate)) +
  geom_hline(yintercept = mean(malarone.data$growth_rate, na.rm = T), linetype = "dashed")+
  geom_boxplot(color ="black", outlier.shape = NA, size=.5, width=.3, na.rm = T)+
  geom_jitter(color='lightgrey', width = 0.2, height = 0, na.rm = T)+
  stat_n_text(na.rm = T, y.pos = -Inf, vjust = -1) +
  labs(x="", y="Std. growth rate")+
  stat_summary(fun=mean, colour = c("#1F78B4", "#33A02C", "#FF7F00"), geom="errorbar", 
               size=1.5 , width=0.05, show.legend = FALSE, 
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x), na.rm = T)+
  stat_summary(fun=mean, colour = c("#1F78B4", "#33A02C", "#FF7F00"), geom="point", 
               size=5 ,show.legend = FALSE, na.rm = T) + 
  theme_article()+
  theme(legend.position="none",
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"))
```

Modelling treatment effects while controlling for age and year-specific variation

```{r}
## Here, scaling is IMPORTANT
lmer_gr_treat <- lmerTest::lmer(growth_rate ~ Treatment + avg_age + days_resample + Sex + Year + (1|Nest),
                                data = malarone.data)

## Summarise model
## -----------------------------------------------------------------------------
summary.model <- summary(lmer_gr_treat)[["coefficients"]]
pander::pandoc.table(summary.model)


## Check model assumptions: okayish
## -----------------------------------------------------------------------------
performance::check_model(lmer_gr_treat)
performance::model_performance(lmer_gr_treat)

## Predicted values for each treatment
## -----------------------------------------------------------------------------
mod_gr <- get_model_data(lmer_gr_treat,type='pred',terms="Treatment") %>% 
   mutate(x = factor(c("Control","Water","Malarone"), levels = c("Control","Water","Malarone")))

## add sample size to labels
## -----------------------------------------------------------------------------
n <- table(malarone.data$Treatment[!is.na(malarone.data$growth_rate)])
mod_gr <- mod_gr %>% 
  mutate(x = paste0(x,"\nn=",n))

g2_pred <- ggplot(mod_gr, aes(x = x, y = predicted)) +
    # geom_hline(yintercept = mean(mod_gr$predicted, na.rm = T), linetype = "dashed") +
    labs(x = "", y = "log growth rate [g/day]\n(z-transformed)")+
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), colour = 'black', size = 1.5, width = .1)+
    geom_errorbar(aes(ymin = predicted - std.error , ymax = predicted + std.error),
                  size = 8, colour = c("#1F78B4", "#33A02C", "#FF7F00"), width = 0)+
    geom_point(color ="black", size = 2)+
    theme_article(base_size = 12)+
    theme(legend.position="none", 
          axis.line.x = element_blank(), 
          axis.ticks.x = element_blank(),
          plot.margin = margin(0, 0, 0, 0, "cm"))  +
  scale_y_continuous(breaks = seq(0, 0.6, 0.1), limits = c(-0.06, 0.64))
```

### `Bloodchemistry`
#### Comparing parameter distributions among groups

First, visualising raw data 

```{r, warning=FALSE, message=FALSE, results='hide'}
## A: Age
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(Age),
            SD = sd(Age),
            LCL = quantile(Age, 0.025),
            UCL = quantile(Age, 0.975))
A <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = Age),
              height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data,
            aes(y = Age, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "A") +
  scale_y_continuous(breaks = seq(15, 40, 5)) +
  ylab("Nestling age\n(days)")

## B: Body condition index
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(BCI),
            SD = sd(BCI),
            LCL = quantile(BCI, 0.025),
            UCL = quantile(BCI, 0.975))
B <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = BCI), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = BCI, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "B") +
  scale_y_continuous(breaks = seq(-100, 200, 50)) +
  ylab("Body condition index\n")

## C: Leucocytozoon
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(Leucocytozoon),
            SD = sd(Leucocytozoon),
            LCL = quantile(Leucocytozoon, 0.025),
            UCL = quantile(Leucocytozoon, 0.975))
C <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = Leucocytozoon), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = Leucocytozoon, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "C") +
  scale_y_continuous(breaks = seq(0, 4, 1)) +
  ylab("Leucocytozoon\ninfection intensity")

## D: AP
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(AP),
            SD = sd(AP),
            LCL = quantile(AP, 0.025),
            UCL = quantile(AP, 0.975))
D <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = AP), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = AP, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "D") +
  ylab("AP\n(U/I)")

## E: AST
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(AST),
            SD = sd(AST),
            LCL = quantile(AST, 0.025),
            UCL = quantile(AST, 0.975))
E <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = AST), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = AST, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "E") +
  ylab("AST\n(U/I)")


## F: GGT
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(GGT),
            SD = sd(GGT),
            LCL = quantile(GGT, 0.025),
            UCL = quantile(GGT, 0.975))
F <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = GGT), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = GGT, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "F") +
  ylab("GGT\n(U/I)")

## G: BuChE
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(BuChE),
            SD = sd(BuChE),
            LCL = quantile(BuChE, 0.025),
            UCL = quantile(BuChE, 0.975))
G <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = BuChE), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = BuChE, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "G") +
  ylab("BuChE\n(KU/I)")

## H: BA
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(BA),
            SD = sd(BA),
            LCL = quantile(BA, 0.025),
            UCL = quantile(BA, 0.975))
H <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = BA), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = BA, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "H") +
  ylab("BA\n(µmol/L)")

## I: LDH
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(LDH),
            SD = sd(LDH),
            LCL = quantile(LDH, 0.025),
            UCL = quantile(LDH, 0.975))
I <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = LDH), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = LDH, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "I") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  ylab("LDH\n(U/I)")

## J: CK
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(CK),
            SD = sd(CK),
            LCL = quantile(CK, 0.025),
            UCL = quantile(CK, 0.975))
J <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = CK), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = CK, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "J") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  ylab("CK\n(U/I)")

## K: ALB
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(ALB),
            SD = sd(ALB),
            LCL = quantile(ALB, 0.025),
            UCL = quantile(ALB, 0.975))
K <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = ALB), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = ALB, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "K") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  ylab("ALB\n(g/L)")

## L: TP
## -----------------------------------------------------------------------------
means <- bloodchem.data %>% 
  group_by(Treatment) %>%
  summarise(MEAN = mean(TP),
            SD = sd(TP),
            LCL = quantile(TP, 0.025),
            UCL = quantile(TP, 0.975))
L <-
  ggplot(means, aes(x = Treatment, y = MEAN)) +
  theme_article(base_size = 10) +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  scale_color_manual(values = c("#A6CEE3", "#1F78B4", "#FDBF6F", "#FF7F00")) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL),
                width = 0.1, size = 1) +
  geom_errorbar(aes(col = Treatment, ymin = MEAN - SD, ymax = MEAN + SD),
                width = 0, size = 8) +
  geom_point(size = 2) +
  geom_jitter(data = bloodchem.data, aes(y = TP), height = 0, size = 0.8, alpha = .6, width = .2, shape = 1) +  
  geom_line(data = bloodchem.data, aes(y = TP, group = Ring), col = "Black", lty = "dotted") +
  labs(tag = "L") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  ylab("TP\n(g/L)")
```


#### Testing for among group differences

Data are not normally distributed, and thus all group comparisons are performed using non-parametric tests, i.e., Wilcoxon signed rank test for paired data and Mann-Whitney test for unpaired data.

The following function is defined to estimate these statistics for all parameters:

```{r}
## function to compute paired and unpaired tests
## -----------------------------------------------------------------------------
u_test <- function(var = NULL, x_name = "Post-Mal", y_name = "Post-Control", 
                   paired = FALSE) {
  
  ## Get data of groups to compare
  ## ---------------------------------------------------------------------------
  x <- bloodchem.data[[var]][bloodchem.data$Treatment == x_name]
  y <- bloodchem.data[[var]][bloodchem.data$Treatment == y_name]
  
  ## Run test and retrun results  
  ## ---------------------------------------------------------------------------
  u <- wilcox.test(x, y, paired = paired)
  
  return(data.frame(parameter = var, 
                    group_x = x_name,
                    group_y = y_name,
                    U_value = u$statistic,
                    p_value = u$p.value))
}
```

Apply the function to all comparisons of interest

```{r, warning=FALSE}
## Unpaired tests 
## -----------------------------------------------------------------------------
u_tests <- rbind(
  ## 1.) Pre-Control vs Pre-Mal, unpaired 
  lapply(names(bloodchem.data)[c(4:7, 9:13)], u_test, 
         x_name = "Pre-Control", y_name = "Pre-Mal") %>% 
    do.call("rbind",.) %>%
    mutate(p_value_adj = p.adjust(p_value, method = "BH")),
  ## 2.) Post-Control vs Post-Mal, unpaired
  lapply(names(bloodchem.data)[c(4:7, 9:13)], u_test, 
         x_name = "Post-Control", y_name = "Post-Mal") %>% 
    do.call("rbind",.) %>%
    mutate(p_value_adj = p.adjust(p_value, method = "BH"))
)
write.csv(u_tests, file = "data/u_tests.csv")
knitr::knit_print(u_tests)

## Paired tests
## -----------------------------------------------------------------------------
w_tests <- lapply(names(bloodchem.data)[c(4:7, 9:13)], u_test, paired = T,
                  x_name = "Pre-Mal", y_name = "Post-Mal") %>% 
  do.call("rbind",.) %>%
  mutate(p_value_adj = p.adjust(p_value, method = "BH"))
write.csv(w_tests, file = "data/w_tests.csv")
knitr::knit_print(w_tests)
```

## Summary
### Model table

```{r, warning=FALSE, message=FALSE}
sjPlot::tab_model(
  model.bci,lmer_gr_treat,
  show.ci50 = T ,
  show.stat = T,
  show.reflvl = T,
  string.stat = "t",
  p.val = "kr",
  dv.labels = c("Std. body condition","Std. growth rate"))
```

### Multiplots

```{r}
## Malarone experiment
## -----------------------------------------------------------------------------
g1_pred_delta + g2_pred + plot_annotation(tag_levels = 'A')& 
  theme(plot.tag = element_text(size = 14))
ggsave(filename = "figures/MalaroneMultiplot.png", device = "png", dpi = 300,
       width = 9, height = 6, units = "in")

## Bloodchemistry data
## -----------------------------------------------------------------------------
A + B + C + D + E + F + G + H + I + J + K + L
ggsave(filename = "figures/BloodChemistryMultiplot.png", device = "png", dpi = 300,
       width = 9, height = 6, units = "in")
```